---
title: "DataCleaningforReanalysisAffiliationChange"
author: "JJ"
date: "October 21, 2014"
output: html_document
---

What I'm trying to do is to make the long data into wide data format so that each ID is a row. 


* import data
```{r}

maindata <- read.csv("./data/pcps.data.csv")
sepVar <- read.csv("./data/SeparationVariableClean.csv")

```

* examine the data
```{r, echo=FALSE}
head(maindata)
names(maindata)

head(sepVar)
names(sepVar)
```

* too many variables. I don't know what I'm going to do with all these.
* step 1: work with the main data. save only useful variables
* variables needed: ID, mat, sex, age, rank, grmini, grmini... [, c(13:21)]
```{r, echo=FALSE}
library(dplyr)
maindata.cleanvar <- 
  select(maindata, ID, Period, Matriline, Sex, Age, Rank, matches("grm|hud"))
head(maindata.cleanvar)
```

* although the df maindata.cleanvar looks better, it is still too much and it requires different process of cleaning. 
* I'll split it into two dfs, with the first one being attributes of IDs.

```{r, echo=FALSE}
attributes <- maindata.cleanvar %>% 
  filter(Period == "PreRemoval") %>%
  select(ID, Matriline, Sex, Age, Rank)

names(attributes)
write.csv(attributes, "./CleanData/attributes.csv", row.names = FALSE)
```

* ok, now the tough ones. first save the aff dataframe
```{r, echo=FALSE}
aff <- select(maindata.cleanvar, ID, Period, matches("grm|hud"))

head(aff)
names(aff)

aff.pre <- aff %>%
  filter(Period == "PreRemoval") %>%
  select(ID, matches("grm|hud")) 
names(aff.pre)
names(aff.pre)[-1] <- paste(names(aff.pre)[-1], "pre", sep = ".")  
names(aff.pre)



aff.post <- aff %>%
  filter(Period == "PostRemoval") %>%
  select(ID, matches("grm|hud"))
names(aff.post)[-1] <- paste(names(aff.post)[-1], "post", sep = ".")  
names(aff.post)
head(aff.post)


aff.clean <- merge(aff.pre, aff.post, by = "ID")
head(aff.clean)
names(aff.clean)

write.csv(aff.clean, "./CleanData/affClean.csv", row.names = FALSE)
```