---
title: "personality models"
author: "JJ"
date: "Thursday, October 23, 2014"
output: html_document
---

* Import Data
```{r, echo=FALSE}
# grmrec.df <- read.csv("./CleanData//grmrec.df.csv")


# hud.df <- read.csv("./CleanData//hud.df.csv")

grmrec.df <- read.csv("./CleanData//grmrec.df.cleancol.csv")
hud.df <- read.csv("./CleanData//hud.df.cleancol.csv")
```

* Center Continuous variables
```{r}
names(grmrec.df)
names(grmrec.df)[c(2,3,4,10, 13:19)]
a <- data.frame(scale(grmrec.df[,c(2,3,4,10, 13:19)], center = TRUE, scale = TRUE))
names(a) <- paste(names(a), "scaled", sep = ".")
grmrec.df.scaled <- cbind(grmrec.df, a)
summary(grmrec.df.scaled)


names(hud.df)
b <- data.frame(scale(hud.df[, c(2:4, 10, 13:19)], center = TRUE, scale = TRUE))
names(b) <- paste(names(b), "scaled", sep = ".")
hud.df.scaled <- cbind(hud.df, b)
summary(hud.df.scaled)
```


# groom Receive in general
```{r, echo=FALSE}
names(grmrec.df)
# null model
grmrec.null.lm <- lm(grmrec.post ~ grmrec.pre,
                        data = grmrec.df)
summary(grmrec.null.lm)
```


```{r}
# personality x pre aff interaction 
grmrec.personality.lm <- lm(grmrec.post ~ grmrec.pre + 
                          ExcitabilityPre  + SociabilityPre
                          + ConfidencePre * grmrec.pre ,
                        data = grmrec.df)
summary(grmrec.personality.lm)
anova(grmrec.personality.lm, grmrec.null.lm)

```

Depending on grmrec.pre, Confidence affect grmrec post differently.


scaled lm model
```{r}
grmrec.personality.scaled.lm <- lm(grmrec.post ~ grmrec.pre.scaled + 
                          ExcitabilityPre.scaled  + SociabilityPre.scaled
                          + ConfidencePre.scaled * grmrec.pre.scaled ,
                        data = grmrec.df.scaled)
summary(grmrec.personality.scaled.lm)
```


## Plot the effects
```{r: plot the Confidence x grmrec.pre}
####New data
grmrec.pre.scaled <- -1:1
SociabilityPre.scaled <- -1:1
ConfidencePre.scaled <- -1:1
ExcitabilityPre.scaled <- -1:1

newgrmrec.personality <- expand.grid(grmrec.pre.scaled, 
                       SociabilityPre.scaled, 
                       ConfidencePre.scaled, 
                       ExcitabilityPre.scaled)
names(newgrmrec.personality) <- c("grmrec.pre.scaled", 
                    "SociabilityPre.scaled", 
                    "ConfidencePre.scaled", 
                    "ExcitabilityPre.scaled")

#### predict y-hat
grmrec.post.hat <- predict(grmrec.personality.scaled.lm, newgrmrec.personality)
newgrmrec.personality.predict <- cbind(grmrec.post.hat, newgrmrec.personality)

#### plot interaction between ConfidencePre and grmrec.pre
grmrec.post.newdata <- aggregate(grmrec.post.hat ~ ConfidencePre.scaled + grmrec.pre.scaled, 
                              newgrmrec.personality.predict, mean)

grmrec.post.newdata$ConfidencePre.scaled <- factor(grmrec.post.newdata$ConfidencePre.scaled,
                                          labels = c("Low Confidence", 
                                                     "Average Confidence", 
                                                     "High Confidence"))
#### plotting
library(ggplot2)
gp <- ggplot(grmrec.post.newdata, 
             aes(x = grmrec.pre.scaled, y = grmrec.post.hat, 
                 colour = ConfidencePre.scaled, group = ConfidencePre.scaled))

grmrec.plot <- gp + geom_line(aes(linetype = ConfidencePre.scaled), size = 1) + theme_bw() + 
  xlab("Levels of Received Grooming (PreRemoval)") +
  ylab("Levels of Received Grooming (PostRemoval)") + 
  theme(text = element_text(size=11), axis.text.x = element_text(vjust=1), 
        legend.position=c(0.20, 0.90), legend.text=element_text(size=10), 
        legend.title=element_text(size=10),
        strip.text = element_text(size = 10)) +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
  ) +
  
  #draws x and y axis line
  theme(axis.line = element_line(color = 'black'))

grmrec.plot
```





# groom receive from nonkin
```{r}
grmrecNonkin.null.lm <- lm(grmrecNonkin.post ~ grmrecNonkin.pre, data = grmrec.df)
summary(grmrecNonkin.null.lm)


names(grmrec.df)
grmrecNonkin.personality.lm <- lm(grmrecNonkin.post ~ grmrecNonkin.pre +
                          ExcitabilityPre * grmrecNonkin.pre +
                          + ConfidencePre  + 
                            SociabilityPre ,
                        data = grmrec.df)
summary(grmrecNonkin.personality.lm)
anova(grmrecNonkin.personality.lm, grmrecNonkin.null.lm)

```

* centered model
```{r}
names(grmrec.df.scaled)
grmrecNonkin.personality.scaled.lm <- lm(grmrecNonkin.post ~ grmrecNonkin.pre.scaled +
                          ExcitabilityPre.scaled * grmrecNonkin.pre.scaled +
                          + ConfidencePre.scaled  + 
                            SociabilityPre.scaled ,
                        data = grmrec.df.scaled)
summary(grmrecNonkin.personality.scaled.lm)
library(car)
sqrt(vif(grmrecNonkin.personality.scaled.lm))
```

* plotting
```{r}
# new data
grmrecNonkin.pre.scaled <- -1:1
ExcitabilityPre.scaled <- -1:1
ConfidencePre.scaled <- -1:1
SociabilityPre.scaled <- -1:1

grmrecNonkin.newdata <- expand.grid(grmrecNonkin.pre.scaled, ExcitabilityPre.scaled,
                       ConfidencePre.scaled, SociabilityPre.scaled)
names(grmrecNonkin.newdata) <- c("grmrecNonkin.pre.scaled", "ExcitabilityPre.scaled", 
                    "ConfidencePre.scaled", 
                    "SociabilityPre.scaled")
grmrecNonkin.post.hat <- predict(grmrecNonkin.personality.scaled.lm, grmrecNonkin.newdata)
grmrecNonkin.newdata.predicted <- cbind(grmrecNonkin.post.hat, grmrecNonkin.newdata)
# aggregate
grmrecNonkin.plotdata <- aggregate(grmrecNonkin.post.hat ~ 
            ExcitabilityPre.scaled + grmrecNonkin.pre.scaled,
          grmrecNonkin.newdata.predicted, mean)
grmrecNonkin.plotdata$ExcitabilityPre.scaled <- 
  factor(grmrecNonkin.plotdata$ExcitabilityPre.scaled, 
       labels = c("Low Excitability", "Average Excitability", "High Excitability"))

gp <- ggplot(grmrecNonkin.plotdata, 
       aes(x = grmrecNonkin.pre.scaled, y = grmrecNonkin.post.hat, 
           colour = ExcitabilityPre.scaled, group = ExcitabilityPre.scaled))
grmrecNonkin.plot <- gp + geom_line(aes(linetype = ExcitabilityPre.scaled), size = 1) + theme_bw() + 
  xlab("Levels of Received Grooming from Kin (PreRemoval)") +
  ylab("Levels of Received Grooming from Kin (PostRemoval)") + 
  theme(text = element_text(size=11), axis.text.x = element_text(vjust=1), 
        legend.position=c(0.20, 0.90), legend.text=element_text(size=10), 
        legend.title=element_text(size=10),
        strip.text = element_text(size = 10)) +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
  ) +
  
  #draws x and y axis line
  theme(axis.line = element_line(color = 'black'))

grmrecNonkin.plot

```


# Huddle

## Huddle in general
```{r, echo=FALSE}
hud.control.lm <- lm(hud.post ~   
                       MatrilineSizePre * hud.pre,
                     data = hud.df)
summary(hud.control.lm)

hud.personality.lm <- lm(hud.post ~ MatrilineSizePre * hud.pre + 
                          ExcitabilityPre * ProportionMatrilineRemoved +
                          ConfidencePre + 
                            SociabilityPre * ProportionMatrilineRemoved,
                        data = hud.df)
summary(hud.personality.lm)
anova(hud.personality.lm, hud.control.lm)
```

* Centered model
```{r}

names(hud.df.scaled)
hud.personality.scaled.lm <- lm(hud.post ~ MatrilineSizePre.scaled * hud.pre.scaled + 
                          ExcitabilityPre.scaled * ProportionMatrilineRemoved.scaled +
                          ConfidencePre.scaled + 
                            SociabilityPre.scaled * ProportionMatrilineRemoved.scaled,
                        data = hud.df.scaled)
summary(hud.personality.scaled.lm)

```

* plotting
```{r}
hud.newdata <- expand.grid(MatrilineSizePre.scaled = -1:1, hud.pre.scaled = -1:1, 
                           ProportionMatrilineRemoved.scaled = -1:1,
                           ExcitabilityPre.scaled = -1:1,
                           ConfidencePre.scaled = -1:1,
                           SociabilityPre.scaled = -1:1)
hud.post.hat <- predict(hud.personality.scaled.lm, hud.newdata)

hud.newdata.predicted <- data.frame(cbind(hud.newdata, hud.post.hat))

hud.exc <- aggregate(hud.post.hat ~ 
                       ProportionMatrilineRemoved.scaled + ExcitabilityPre.scaled,
          hud.newdata.predicted, mean)
factor(hud.exc$ExcitabilityPre.scaled)
hud.exc$ExcitabilityPre.scaled <- factor(hud.exc$ExcitabilityPre.scaled, 
                                         labels = c("Low Excitability", 
                                                  "Average Excitability",
                                                  "High Excitability"))

hud.soc <- aggregate(hud.post.hat ~ 
                       ProportionMatrilineRemoved.scaled + SociabilityPre.scaled,
                     hud.newdata.predicted, mean)
hud.soc$SociabilityPre.scaled <- factor(hud.soc$SociabilityPre.scaled,
                                        labels = c("Low Sociability", 
                                                  "Average Sociability",
                                                  "High Sociability"))
```

* plot Excitability
```{r}
gp <- ggplot(hud.exc, 
             aes(x = ProportionMatrilineRemoved.scaled, y = hud.post.hat, 
                 group = ExcitabilityPre.scaled, colour = ExcitabilityPre.scaled))

hud.exc <- gp + geom_line(aes(linetype = ExcitabilityPre.scaled), size = 1) + theme_bw() + 
  
  # change axis labels here
  
  xlab("Proportion of Matriline being Removed") +
  ylab("Levels of huddling (PostRemoval)") + 
  
  theme(text = element_text(size=11), axis.text.x = element_text(vjust=1), 
        legend.position=c(0.20, 0.90), legend.text=element_text(size=10), 
        legend.title=element_text(size=10),
        strip.text = element_text(size = 10)) +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
  ) +
  
  #draws x and y axis line
  theme(axis.line = element_line(color = 'black'))

hud.exc
```

* plot Sociability
```{r}
gp <- ggplot(hud.soc, 
             aes(x = ProportionMatrilineRemoved.scaled, y = hud.post.hat, 
                 group = SociabilityPre.scaled, colour = SociabilityPre.scaled))

hud.soc <- gp + geom_line(aes(linetype = SociabilityPre.scaled), size = 1) + theme_bw() + 
  
  # change axis labels here
  
  xlab("Proportion of Matriline being Removed") +
  ylab("Levels of huddling (PostRemoval)") + 
  
  theme(text = element_text(size=11), axis.text.x = element_text(vjust=1), 
        legend.position=c(0.90, 0.15), legend.text=element_text(size=10), 
        legend.title=element_text(size=10),
        strip.text = element_text(size = 10)) +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
  ) +
  
  #draws x and y axis line
  theme(axis.line = element_line(color = 'black'))

hud.soc

```


## Huddle with Kin
```{r, echo=FALSE}

hudKin.control.lm <- lm(hudKin.post ~ hudKin.pre * Age +
                          hudKin.pre * MatrilineSizePre, data = hud.df)
summary(hudKin.control.lm)

hudKin.personality.lm <- lm(hudKin.post ~ hudKin.pre * Age +
                      hudKin.pre * MatrilineSizePre + 
                          ExcitabilityPre * ProportionMatrilineRemoved +
                          + ConfidencePre + 
                            SociabilityPre,
                        data = hud.df)
summary(hudKin.personality.lm)
anova(hudKin.personality.lm, hudKin.control.lm)

```

* centered model
```{r}
hudkin.personality.scaled.lm <- lm(hudKin.post ~ hudKin.pre.scaled * Age.scaled +
                      hudKin.pre.scaled * MatrilineSizePre.scaled + 
                          ExcitabilityPre.scaled * ProportionMatrilineRemoved.scaled +
                          + ConfidencePre.scaled + 
                            SociabilityPre.scaled,
                        data = hud.df.scaled)
summary(hudkin.personality.scaled.lm)
```

* plot Exc x separation
```{r}
hudkin.newdata <- expand.grid(hudKin.pre.scaled = -1:1, Age.scaled = -1:1,
                              MatrilineSizePre.scaled = -1:1, 
                              ProportionMatrilineRemoved.scaled = -1:1,
                              ExcitabilityPre.scaled = -1:1,
                              ConfidencePre.scaled = -1:1,
                              SociabilityPre.scaled = -1:1)

hudKin.post.hat <- predict(hudkin.personality.scaled.lm, hudkin.newdata)
hudkin.newdata.predicted <- data.frame(cbind(hudKin.post.hat, hudkin.newdata))
hudkin.plot.data <- 
  aggregate(hudKin.post.hat ~ ExcitabilityPre.scaled + ProportionMatrilineRemoved.scaled,
          hudkin.newdata.predicted, mean)
hudkin.plot.data$ExcitabilityPre.scaled <- 
  factor(hudkin.plot.data$ExcitabilityPre.scaled, 
         labels = c("Low Excitability", "Average Excitability", "High Excitability"))
names(hudkin.plot.data)
gp <- ggplot(hudkin.plot.data, 
       aes(x = ProportionMatrilineRemoved.scaled, y = hudKin.post.hat, 
           group = ExcitabilityPre.scaled, colour = ExcitabilityPre.scaled))
hudKin.plot <- gp + geom_line(aes(linetype = ExcitabilityPre.scaled), size = 1) + 
  theme_bw() + 
  
  # change axis labels here
  
  xlab("Proportion of Matriline being Removed") +
  ylab("Levels of huddling with Kin (PostRemoval)") + 
  
  theme(text = element_text(size=11), axis.text.x = element_text(vjust=1), 
        legend.position=c(0.20, 0.85), legend.text=element_text(size=10), 
        legend.title=element_text(size=10),
        strip.text = element_text(size = 10)) +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
  ) +
  
  #draws x and y axis line
  theme(axis.line = element_line(color = 'black'))
hudKin.plot

```
